---
title: "Site Frequency Spectrum"
author: "Bruce Cochrane"
date: "1/5/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library (tidyverse)
```

II Site frequency spectrum

From sfsDNA.Rmd.  Will use our sfs function with plot equals false; then plot the results with ggplot.  (Also need to see what we can do with the PEGAS function)

```{r}
set.seed(123)  
rand <-ms(nsam=10, nreps=1,opts="-s 10")  
rand <-read.ms.output(txt=rand)  
x <-rand$gametes[[1]]  
x  
```

So that gives the gametic array, to which we can apply sfs:

```{r}
sf <- sfs(x,pl=FALSE)
sf
```

Not sure how we want to do this in the long run, but making a data frame works for now
```{r}
sf <- data.frame(i=1:length(sf[[2]]),Frequency=sf[[2]])
sf
```

```{r}
gg.spectrum <- function(sf,unfolded=FALSE){
sf <- data.frame(i=1:length(sf),Frequency=sf)
ttl <- ifelse(unfolded,"Unfolded Site Frequency Spectrum","Folded Site Frequency Spectrum")
plt <- ggplot(sf,aes(i,Frequency)) +
         geom_col(alpha=.4,fill="red") +
        ggtitle(ttl) +
        scale_x_continuous(breaks=c(1:nrow(sf))) +
        scale_y_continuous(breaks=c(1:max(sf$Frequency)))

plt
}
```

Probably can add some more stuff to get rid of subticks.  


OK.  So that should be an easy functional change (if we decide to go that route.  Now let's look and see if we can do something similar with Pegas function site.spectrum()

```{r}
acp29 <-read.FASTA("./Data/acp29.fasta")
sp2 <- site.spectrum(acp29)
sp3 <- as.numeric(sp2)
```

OK.  Try the above function on sp2 as is:

```{r}
gg.spectrum(sp3)

```

OK.  Good to go; eventually need to add faceting.

