---
title: "Trees"
author: "Bruce Cochrane"
date: "1/5/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggtree)
```

install chunk

```{r,eval=FALSE}
source("https://bioconductor.org/biocLite.R")
biocLite("ggtree")
```


This is the big one

## Single tree from ms

```{r}
tr <-ms(nsam=10,nreps=1,opts=c("-t 10 -T"))
tr2 <-read.tree(text=tr)
plot(tr2)
```

```{r}
ggtree(tr2) +
  geom_text2(aes(subset=!isTip, label=node), hjust=-.3,col="red",size=3) +
  geom_tiplab() +
  geom_nodepoint()
```

Bottom line - probably don't need to change the functions here for basic ms-type stuff, however for building pretty trees, it's a good one
## Multiple trees

Here's the basic (stolen) code for doing a multitree

```{r}
trees <- lapply(c(10, 20, 40,25), rtree)
class(trees) <- "multiPhylo"
ggtree(trees) + facet_wrap(~.id, scale="free",ncol=2) + geom_tiplab(size=2)
```

So the only trick is to add the facet_wrap; the scale allows each tree to be proportioned independently.
## Multiple panels with different graphics