---
title: <h3>Testing RStudio New Tricks</h3>
author: "Bruce Cochrane"
date: "12/11/2016"
output: 
  html_document: 
    fig_caption: yes
    number_sections: yes
    theme: paper
    toc: yes
    code_folding: show
---
# Overview {.tabset .tabset-fade .tabset-pill}

This is just a foolaround file with the new RStudio shortcuts (available at file:///Applications/RStudio.app/Contents/Resources/www/docs/keyboard.htm).  

1.  Assignment - supposed to be option -, however only works in chunk
2.  pipe operator - CmdSftM -- %>% -- works fine
3.  output yaml can be modified by clicking on the gear wheel at the top of the editing pane, followed by output options
4.  Note - under global options/rmarkdown, uncheck show chunk output inline to retain traditional RSTudio look 

## Some more detail

So I wonder if we could do this with the maps project - interesting possibilities

## Some different material

Put in at this point just to see what a subheading would look like

# Some basic analysis {.tabset}

First, we will generate some numbers, then we will plot it in a couple of ways:

```{r}
x <- rnorm(100,0,1)
```

By clicking on the tabs below, you can see the data visualized as either a conventional histogram, or as a "qqplot", which is a good way to determine the extent to which observed data are explained by a normal distribution.

## A histogram

```{r,fig.cap="Figure 1.  A Histogram of 100 Normally Distributed numbers with mean of zero and standard deviation of one"}
hist(x)
```

## A qqplot
```{r,fig.cap="Figure 2.  QQPlot, comparing the distribution of the data to that expected based on a normal distribution"}
qqnorm(x)
```

---

So with the chapters, we need to work on the yaml for the master file and then add tabsets within each as needed.